---
title: "homework chapter 25"
author: "xiang liu"
date: "April 31st, 2015"
output:
  html_document:
    fig_height: 4
    fig_width: 8
    toc: yes
  pdf_document:
    fig_height: 4.3
    fig_width: 8
    keep_tex: yes
    toc: yes
geometry: margin=1in
setspace: onehalfspacing
fontsize: 12pt
---


**25.9(a)**

\hfill

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE}
library(pander)
library(doBy)
data=read.table("CH25PR09.txt",header=FALSE, 
                col.names = c('characteristics', 'machine', 'coil'))
attach(data)
machine <- factor(machine)
fit <- lm(characteristics ~ machine)
pander(anova(fit))
```

* Ho: $\sigma^2_\mu$ = 0 vs. Ha: $\sigma^2_\mu > 0$
* Decision Rule: since the test statistics F(3,36) = 28.1, p < 0.1, we reject the Ho.
* Conclusion: there is sufficient evidence to suggest all machinese in the plant produced different mean coil characteristic.

\hfill

**25.10(d)**

* $\hat \sigma^2_\mu$ = (MSTR - MSE)/n = (200.8-7.15)/10 = 19.37

\hfill

**25.15(a)** 

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE}
data=read.table("CH25PR15.txt",header=FALSE, 
                col.names = c('MPG', 'driver', 'car','n'))
attach(data)
driver <- factor(driver)
car <- factor(car)
fit <- lm(MPG ~ driver + car + driver:car)
pander(anova(fit))
```

* Ho: $\sigma^2_{\alpha \beta}$ = 0 vs. Ha: $\sigma^2_{\alpha \beta} > 0$
* Decision Rule: since the test statistics F(12,20) = 2.45, p = 0.3715 > 0.05, we cannot reject the Ho.
* Conclusion: there is insufficient evidence to suggest an interaction effect between divers and cars on MPG. 

\hfill

**25.15(c)** 

* MSA = 93.43, MSB = 23.68, MSAB = 0.2049, n = 2, a = 4, b = 5
* $\hat \sigma^2_a$ = (MSA - MSAB)/bn = 9.3225
* $\hat \sigma^2_b$ = (MSB - MSAB)/an = 2.9344
* Factor A, driver, seem to have a greater effect on MPG.
