---
title: "homework chapter 19"
author: "xiang liu"
date: "March 8th, 2015"
output:
  html_document:
    fig_height: 4
    fig_width: 8
    toc: yes
  pdf_document:
    fig_height: 4.3
    fig_width: 8
    keep_tex: yes
    toc: yes
geometry: margin=1in
setspace: onehalfspacing
fontsize: 22pt
---

**Based on KNNL Problems 19.16 & 19.17 on Page 868-869**. Disk drive service.The staff of a service center for electronic equipment includes three technicians who specialize in repairing three widely used makes of disk drives for desktop computers. It was desired to study the effects of technician (factor A) and make of disk drive (factor B) on the service time. Data was collected on the number of minutes required to complete the repair job in a study where each technician was randomly assigned to five jobs on each make of the disk drive. The data are in file “CH19PR16.txt”.

\hfill

**(a)**  Propose a two-way ANOVA model with interactions. Discuss the assumptions and the conditions of the model.

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE}
library(pander)
library(doBy)
data.19=read.table("CH19PR16.txt",col.names=c('Y','A','B','i'))
names(data.19)=c("time","technician","make",'id')
data.19$technician = factor(data.19$technician)
data.19$make = factor(data.19$make)
```
* Assumption #1: Dependent variable has to be measured at the continuous level
* Assumption #2: Explanatory variables are independent (each has more than one level). 
* Assumption #3: Independence of observations
* Assumption #4: Equal variances for each combination of the groups of the two independent variables
* Assumption #5: Dependent variable should be approximately normally distributed for each combination of the groups of the two independent variables

\hfill

**(b)** Create an interaction plot and comment on it.

\hfill

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE}
library(lattice)
stripplot(time~make, groups=technician, auto.key=list(space='right', text = c("mode1", "mode2", "mode3"), points=F,lines=T), points = T, jitter.data=T, type=c("p","a"), data=data.19, xlab="Technician", ylab="Time")
# Another way to draw interaction plot
# interaction.plot(x.factor = data.19$technician, trace.factor = data.19$make,
#                  response = data.19$time,fun = mean, legend = T, type = "l", 
#                  lwd = 2, ylim = c(30, 80),axes = F, trace.label = "Race", 
#                  xlab = "Sex", ylab = "Clock Drawing Total Score")
# title(main = "IPV History")
# axis(1, at = 1:3)
# axis(2, at = seq(30, 80, 5))
```

`Interpret`: __There appears to be a significant interaction between technician and make of disk drive. As technician 1 is better at repairing mode2, technician 2 is good at repairing mode 1 and technician 3 excells at mode 3__

\hfill

\hfill

**(c)** Estimate $\mu_{22}$ and $\alpha \beta_{22}$

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.align='left'}
a=tapply(data.19$time, INDEX=list(data.19$technician, data.19$make), FUN=mean)
a=cbind(a,rowMeans(a))
a=rbind(a,colMeans(a))
rownames(a)= c('a1','a2','a3','mean(a)')
colnames(a)= c('b1','b2','b3','mean(b)')
print(a,digits=4)
```

`Interpret`: __The $\mu_{22}$ is 61.2. Since $\mu_{ij}=\mu+\alpha_i+\beta_j+\alpha \beta_{ij}$. $\alpha_2$ = 55.27(group mean) - 55.82(grand mean) = -0.55. And $\beta_2$ = 56.6 - 55.82 = 0.78. Therefore, $\alpha \beta_{22}$ = 61.2 - 55.82 - (-0.55) - 0.78 = 5.15.__

\hfill

\hfill

**(d)** Test whether or not the two factors interact. Use $\alpha$ = 0.05. State the hypotheses, decision rule, and conclusion.
```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.align='left'}
panderOptions('digits' , 4)
model = lm(time ~ technician * make, data.19) 
pander(anova(model))
```

* Hypothesis: $H_0$: All of $\alpha \beta_{ij}$ equal to 0 vs $H_a$: not all $\alpha \beta_{ij}$ = 0.
* Decision rule: If p-value corresponding to interaction term is less than 0.05, we reject the null hypothesis. Since p-value < .05, we have sufficient evidence to reject the null hypothesis.
* Conclusion: There is a significant interaction effect between technician and make of disk drive. 

\hfill

\hfill

**(e)** Test whether or not main effects for technician and make of the drive are present. Use $\alpha$ = 0.05. In each case state the hypotheses, decision rule, and conclusion.

* Main effect for Technician
    + Hypothesis: $H_0$: $\alpha_1$ = $\alpha_2$ = $\alpha_3$ = 0 vs $H_a$: not all $\alpha_i$ = 0.
    + Decision rule: If p-value corresponding to technician term is less than 0.05, we reject the null hypothesis. Since p-value > .05, we don't have sufficient evidence to reject the null hypothesis.
    + Conclusion: There isn't a significant main effect of technician. 
  
* Main effect for Make of Disk Driver
    + Hypothesis: $H_0$: $\beta_1$ = $\beta_2$ = $\beta_3$ = 0 vs $H_a$: not all $\beta_j$ = 0.
    + Decision rule: If p-value corresponding to make term is less than 0.05, we reject the null hypothesis. Since p-value > .05, we don't have sufficient evidence to reject the null hypothesis.
    + Conclusion: There isn't a significant main effect of make of disk driver.

\hfill

\hfill

**KNNL Problem 19.4 on Page 864**. In a two-factor study, the treatment means  $\mu_{ij}$ are as follows:

\hfill

**(a)**  Obtain the factor A level means.

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.align='left'}
b=matrix(c(34,23,36,40,29,42),2,3,byrow = T)
b=cbind(b,rowMeans(b))
b=rbind(b,colMeans(b))
colnames(b)= c('B1','B2','B3','mean(B)')
rownames(b)= c('A1','A2','mean(A)')
print(b,digits=4)
```

`Solution`: __As shown in the table, the factor level mean, A1 = 31 and A2 = 37.__

\hfill

**(b)**  Obtain the main effects of factor A

`Solution`: __The main effects of factor A $\alpha_1$ = 31 ($\mu_1$) - 34 (grand mean) = -3 and $\alpha_2$ = 37 ($\mu_2$) - 34 (grand mean) = 3.__

\hfill

**(c)**  Does the fact that $\mu_{12} - \mu_{11}$ = -11 while $\mu_{13} - \mu_{12}$ = 13 imply that factors A and B interact? Explain

`Solution`: __No, these cell mean differences only refer to A1. It only examines the differences within A1 level.__

\hfill

**(d)**  Prepare a treatment means that plots and determine whether the two factors interact. What do you find?

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.height=3}
c=as.data.frame(as.table(b[1:2,1:3]))
names(c) = c("A","B","value")
xyplot(value~B, groups=A, auto.key=list(lines=T, points= F, corner = c(1, 1), cex=1.0), jitter.data=T, type=c("p","a"), data=c)
```

`Interpretation`: __Two lines are parallel so there is no interaction presented__.






