---
title: "homework chapter 18"
author: "xiang liu"
date: "March 1st, 2015"
output:
  html_document:
    fig_height: 4
    fig_width: 8
    toc: yes
  pdf_document:
    fig_height: 4.3
    fig_width: 8
    keep_tex: yes
    toc: yes
geometry: margin=1in
setspace: onehalfspacing
fontsize: 22pt
---


**Filling machines**. A company uses six filling machines of the same make and model to place detergent into cartons that show a label weight of 32 ounces. The production manager has complained that the six machines do not place the same amount of fill into the cartons. A consultant requested that 20 filled cartons be randomly selected from each of the 6 machines and the content of each carton be carefully weighed. The deviations from 32 ounces were recorded. Sample data are stored in the data file “CH16PR11.txt” with variable names $y$, $i$, and $j$, respectively.

Assume that ANOVA model $Y_{ij} = \mu_i + \epsilon_{ij}$ is applicable. In HW3, the model was fitted.

\hfill

**(a)** Obtain a 95% confidence interval for *D* = $\mu_2 - \mu_1$. Interpret your interval estimate.
```{r, echo=T,message=FALSE,comment=NA, warning=FALSE}
library(pander)
library(doBy)
data.18=read.table("CH16PR11.txt",col.names=c('y','i','j'))
names(data.18)=c("amount","machine","replication")
#### statistics table
ft = summaryBy(amount~machine, data.18, FUN=c(length, mean, sd))
attributes(ft)$names <- c("models","sample size",'mean','standard deviation')
pander(ft)
#### the confidence interval
x = subset(data.18,machine==2,amount) - subset(data.18,machine==1,amount)
x = as.matrix(x)
n = length(x)
error = round(qt(0.975,df=n-1)*sd(x)/sqrt(n),3)
left  = mean(x)-error
right = mean(x)+error
print(paste("The 95% confidence interval is between", left, " and ", right))
## machine is the factor in this experiment and has to be declared as such in R
label=c("machine","replication")
data.18[label]<-lapply(data.18[label],factor)
```

`Interpret`: __The detergent placed into carton by the 2nd model is expected to be between  -0.015  and  0.249 higher than it's placed by the 1st model. Since 0 is included in 95% confidence interval, we don't have enough confidence to include significant differences between the first two models.__


\hfill

\hfill

**(b)** Use the Tukey multiple comparison method to determine which pairs of machines differ significantly.

```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.align='left'}
cat("First we noted that there is difference among the 6 models overall")
panderOptions('round', 3)
data.aov=aov(amount ~ machine,data=data.18)
pander(data.aov)
mcTukey=TukeyHSD(data.aov,"machine")$machine
pander(mcTukey)
```

`Interpret`: __Using Tukey comparison method, there is a significant differences between model 1 and 3, model 1 and 4, model 2 and 3, model 2 and 4, model 3 and 5, model 3 and 6, model 4 and 5, and model 4 and 6.__

\hfill

\hfill

**(c)** Machines 1 and 2 were bought new five years ago, machines 3 and 4 were bought in a reconditioned state five years ago, and machines 5 and 6 were bought new last year. Before doing the study, it was decided that the comparison of interest was to compare the average of the mean weight deviations for machines 1 and 2 with the average of the mean weight deviations for machines 3 and 4. Set up an appropriate contrast and estimate it with a 95% confidence interval.
```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.align='left'}
data.18$level=data.18$machine
levels(data.18$level) <- list(A=c(1,2), B=c(3,4),C=c(5,6))
attach(data.18)
```

\hfill

`Interpret`: __Since it's a planned test, Bonferroni procedure is recommended. The Bonferroni adjusted p-value is 1.496721e-09. Therefore, we reject the null hypothesis and conclude that we have 95% confidence there is signifincant difference between the average of the mean weight deviations for model 1 and 2 and that for model 3 and 4.__

\hfill

\hfill

**(d)** Perform a Levene’s test for constant variance assumption.
```{r, echo=T,message=FALSE,comment=NA, warning=FALSE,fig.align='left'}
library(car)
pander(leveneTest(amount~machine, data=data.18))
```

`Interpret`: __Since the p-value is larger than .05, we cannot reject the null hypothesis of equal variance. __
